Leveraging Altruistic Peers to Reduce the Bandwidth Costs of Free Content Downloads

### 2008-06-30 docs/abstract/abstract.tex
Final version of abstract.

\title{Leveraging Altruistic Peers to Reduce the Bandwidth Costs of Free Content Downloads}
\numberofauthors{2}
\author{
\alignauthor Cameron Dale\\
	\affaddr{School of Computing Science (student)}\\
	\affaddr{Simon Fraser University}\\
	\affaddr{Burnaby, British Columbia, Canada}\\
	\email{Email: camerond@cs.sfu.ca}
%\and
\alignauthor Jiangchuan Liu\\
	\affaddr{School of Computing Science (faculty)}\\
	\affaddr{Simon Fraser University}\\
	\affaddr{Burnaby, British Columbia, Canada}\\
	\email{Email: jcliu@cs.sfu.ca}
}

@@

\begin{abstract}
We introduce the opportunity many free content distributors have to
reduce their bandwidth costs using a peer-to-peer download
mechanism. Our model for this peer-to-peer distribution is simple to
implement, backwards-compatible with the existing infrastructure,
and can be deployed incrementally. It makes wide use of a customized
distributed hash table (DHT) to facilitate the finding of peers and
the efficient downloading of files. We also present the example
solution we created for one of these distributors, the Debian
project, which is already in use by some users.
\end{abstract}

@@

There are a large number of free content distributors using custom
package distribution systems over the Internet to distribute content
to their users: most Linux distributions (e.g. Debian, Red Hat,
Gentoo), Cygwin, CPAN, etc.
These distributors have developed many different
methods for this distribution, but they almost exclusively use a
client-server model to satisfy user requests. The contents' size,
popularity and number of users results in a large number of
requests, which usually requires a complicated network of mirrors to handle. Due
to the free nature of this content, many users are willing and able
to contribute upload bandwidth to this distribution, but currently
have no way to do this.

@@

We also present a sample implementation based on the Debian package
distribution system. This implementation is simple, makes use of
mostly well-known technologies, and so serves as an example for
other free content distributors of the opportunity that can be easily met
with such a system.

@@

Most Linux distributions use a software package management system
that fetches packages to be installed from a network of mirrors.
Debian-based distributions uses the \texttt{apt} program, which
downloads Debian packages from one of many mirrors. RPM-based
distributions use \texttt{yum}, and Gentoo uses \texttt{portage},
which both operate in a similar way. Other free software
distributors also use a similar system: CPAN distributes files
containing software packages for the PERL programming language,
using SOAP RPC requests to find and download files; Cygwin provides
many of the standard Unix/Linux tools in a Windows environment,
using a package management tool that requests \texttt{tar} files from
websites; two software distribution systems exist for Mac OSX,
fink and MacPorts, that also retrieve packages in this way. Also,
some systems use direct web downloads, but with a hash verification
file also available. These
hash files usually have the same file name, but with an added
extension identifying the hash used (e.g. \texttt{.md5} for the MD5
hash). Finally, there are other programs that make use of
cryptographic hashing to identify files: e.g. Git is a version
control system in which all files and changes are identified by
their SHA1 hash.

@@

This situation presents a clear opportunity to use a new
peer-to-peer file sharing protocol.
A downloading peer can lookup the package hash using the protocol and, if it
is found, download the file from those peers and verify the content.
If the package hash can not be found, the peer will
fallback to downloading from the original content location, and once
complete will announce to other peers indicating that it now has
the content. The original servers or mirrors thus act as \emph{seeds} for the
P2P system without any modification to them, and users are satisfied even when
there are no peers which allows the system to be deployed incrementally.

>>>>

\title{Leveraging Altruistic Peers to Reduce the Bandwidth Costs of Free Content Downloads}
\numberofauthors{2}
\author{
\alignauthor Cameron Dale\\
	\affaddr{School of Computing Science}\\
	\affaddr{Simon Fraser University}\\
	\affaddr{Burnaby, British Columbia, Canada}\\
	\email{Email: camerond@cs.sfu.ca}
%\and
\alignauthor Jiangchuan Liu\\
	\affaddr{School of Computing Science}\\
	\affaddr{Simon Fraser University}\\
	\affaddr{Burnaby, British Columbia, Canada}\\
	\email{Email: jcliu@cs.sfu.ca}
}

@@

\begin{abstract}
We introduce the opportunity many free content distributors have to
reduce their bandwidth costs using a peer-to-peer download
mechanism. Our model for this peer-to-peer distribution is simple to
implement, backwards-compatible with the existing infrastructure,
and can be deployed incrementally. It makes wide use of a customized
distributed hash table (DHT) to facilitate the finding of peers and
the efficient downloading of files. We also present the example
solution we created for one of these distributors, the Debian
project, which is already seeing some use.
\end{abstract}

@@

There are a large number of free content distributors using custom
package distribution systems over the Internet to distribute content
to their users: e.g. most Linux distributions (Debian, Red Hat,
Gentoo), Cygwin, CPAN, etc.
These distributors have developed many different
methods for this distribution, but they almost exclusively use a
client-server model to satisfy user requests. The contents' size,
popularity and the number of users results in a large number of
requests, which usually requires a complicated network of mirrors to handle. Due
to the free nature of this content, many users are willing
to contribute upload bandwidth to this distribution, but currently
have no way to do this.

@@

We also present a sample implementation based on the Debian package
distribution system. This implementation is simple and makes use of
well-known technologies, and so serves as an example for
other free content distributors of the opportunity that can be easily met
with such a system.

@@

Most Linux distributions use a software package management system
that fetches packages from a network of mirrors.
Debian-based distributions use the \texttt{apt} program, which
downloads Debian packages from one of many mirrors. RPM-based
distributions use \texttt{yum}, and Gentoo uses \texttt{portage},
which both operate in a similar way. Other free software
distributors also use a similar system: CPAN distributes files
containing software packages for the PERL programming language,
using SOAP RPC requests to find and download files; Cygwin provides
many of the standard Unix/Linux tools in a Windows environment,
using a package management tool that requests \texttt{tar} files from
websites; two software distribution systems exist for Mac OSX,
fink and MacPorts, that also retrieve packages in this way. Also,
some systems use direct web downloads, but with a hash verification
file also available. These
hash files usually have the same file name, but with an added
extension identifying the hash used (e.g. \texttt{.md5} for the MD5
hash). Finally, there are other programs that make use of
cryptographic hashing to identify files: e.g. Git is a version
control system in which all files and changes are identified by
their SHA1 hash.

@@

This situation presents a clear opportunity to use a new
peer-to-peer file sharing protocol.
A downloading peer can lookup the package hash using the protocol and, if it
is found, download the file from those peers and verify the content.
If the package hash can not be found, the peer will
fallback to downloading from the original content location, and once
complete will announce to other peers indicating that it now has
the content. The original servers or mirrors thus act as \emph{seeds} for the
P2P system, without any modification to them. Users are satisfied even when
there are no peers, allowing the system to be deployed incrementally.
